---
title: "Feeling Connected"
output:
  distill::distill_article:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

```{r echo=F}
library(tidyverse)
library(ggthemes)
library(plotly)
```

## Overview

Stilesians were asked: 

> _On a scale of 1-10: On average, how connected have you felt to other people in the Stiles community during each time period? (1 = very disconnected, 10 = very connected.) Just answer to your best estimation/memory. First-years, it's ok if your answers are very low for the periods before Fall semester :)._

### Guiding Questions:

* For each year/grade: which time represents the highest connectedness?
* How does connectedness vary between location and enrollment in F20?
* What happens to connectedness over Spring break and Summer?

Also note that the remainder of the year, from November 21 through December, is just assumed to be constant.

### Read Data

```{r}
# Load data
t <- "20201227_235052"
data.connectedness <- read.csv(paste("data/connectedness_", t, ".csv", sep="")) %>%
    mutate(Housing.F20=factor(
        Housing.F20,
        levels=c("On-campus Stiles", "On-campus OC", "Off-campus", "Remote"))) %>%
        #levels=c("Stiles", "Old Campus", "Off-Campus", "Remote"))) %>%
    mutate(Enrolled.F20=ifelse(is.na(Enrolled.F20), NA, ifelse(
        Enrolled.F20==T, "Enrolled", "Not Enrolled"))) %>%
    mutate(Enrolled.F20=factor(
        Enrolled.F20, levels=c("Not Enrolled", "Enrolled")
    )) %>%
    mutate(Year=as.factor(Year))
# "Jitter" values just a little bit, so it's easier to see overlapping points
c_jitter <- data.connectedness %>%
    mutate(Day=jitter(Day, factor=1)) %>%  # x
    mutate(Connectedness=jitter(Connectedness, factor=.5))  # y
```

## First Look: Small Multiples

```{r layout="l-page", fig.cap="Vertical columns represent enrollment status in Fall 2020, horizontal rows represent proximity to campus in Fall 2020."}
g_all <- (
    ggplot(c_jitter %>% filter(!is.na(Housing.F20) & !is.na(Enrolled.F20) & !is.na(Year)),
           aes(x=Day, y=Connectedness, color=Year, group=ID)) +
    geom_line(alpha=0.5) +
    geom_point(alpha=0.6) +
    scale_y_continuous(breaks=1:10) +
    scale_x_continuous(breaks=unique(data.connectedness$Day)) +
    scale_color_colorblind() +
    facet_grid(Housing.F20 ~ Enrolled.F20) +
    theme_minimal()
) %>% ggplotly()
g_all
```

Note that you can selectively view graduation years, or groups of graduation years, by clicking on the legend above.

Focus on the Fall semester of 2020: what's going on in Day 327?

Timeline review:

| Epoch (`t=`value) | Time frame | Date range (2020) | Day of Year (Starting) |
|-------------------|------------|-------------------|------------|
| `t1` | Before Spring Break | < 3/7 | 13 |
| `t2` | Spring Break | 3/7 - 3/22 | 67 |
| `t3` | Remainder of the Spring semester| 3/23 - 5/6 | 83 |
| `t4` | Summer | 5/7 - Late August | 128 |
| `t5` | Fall semester, before Break | 8/31 - 11/21 | 244 |
| `t6` | Remainder of the Fall semester | 11/22 - 2020 | 327 |

## Deep Dives

On summary statistics (means and standard errors).

### Graduation Year

```{r}
data.summ.yr <- data.connectedness %>%
    filter(!is.na(Housing.F20) & !is.na(Enrolled.F20) & !is.na(Year)) %>%
    group_by(Year, Day) %>%
    summarize(mean_se(Connectedness), N..Responses=n()) %>%
    mutate(Connectedness=y)

dodge <- position_dodge(12)
g <- ggplot(data.summ.yr, aes(x=Day, y=Connectedness, color=Year, text=N..Responses)) +
    geom_line(position=dodge) +
    geom_errorbar(aes(ymin=ymin, ymax=ymax), position=dodge, width=10) +
    scale_y_continuous(breaks=1:10) +
    scale_x_continuous(breaks=unique(data.connectedness$Day)) +
    scale_color_colorblind() +
    theme_minimal()
ggplotly(g)
```

Note that setting tooltips with geom_line is funky.^[https://github.com/ropensci/plotly/issues/1153]

### Proximity to Campus

```{r}
data.summ.yr <- data.connectedness %>%
    filter(!is.na(Housing.F20) & !is.na(Enrolled.F20) & !is.na(Year)) %>%
    group_by(Housing.F20, Day) %>%
    summarize(mean_se(Connectedness), N..Responses=n()) %>%
    mutate(Connectedness=y)

dodge <- position_dodge(12)
g <- ggplot(data.summ.yr, aes(x=Day, y=Connectedness, color=Housing.F20, text=N..Responses)) +
    geom_line(position=dodge) +
    geom_errorbar(aes(ymin=ymin, ymax=ymax), position=dodge, width=10) +
    scale_y_continuous(breaks=1:10) +
    scale_x_continuous(breaks=unique(data.connectedness$Day)) +
    scale_color_colorblind() +
    theme_minimal()
ggplotly(g)
```

### Enrollment Status

```{r}
data.summ.yr <- data.connectedness %>%
    filter(!is.na(Housing.F20) & !is.na(Enrolled.F20) & !is.na(Year)) %>%
    group_by(Enrolled.F20, Day) %>%
    summarize(mean_se(Connectedness), N..Responses=n()) %>%
    mutate(Connectedness=y)

dodge <- position_dodge(12)
g <- ggplot(data.summ.yr, aes(x=Day, y=Connectedness, color=Enrolled.F20, text=N..Responses)) +
    geom_line(position=dodge) +
    geom_errorbar(aes(ymin=ymin, ymax=ymax), position=dodge, width=10) +
    scale_y_continuous(breaks=1:10) +
    scale_x_continuous(breaks=unique(data.connectedness$Day)) +
    scale_color_colorblind() +
    theme_minimal()
ggplotly(g)
```


## Granular Views

The following plots offer a more granular look at Stiles-connectivity through time from the perspective of each of the following factors: graduation year, proximity to campus in the Fall, and enrollment status in the Fall.

```{r}
g_year <- (
    ggplot(c_jitter %>% filter(!is.na(Year)),
           aes(x=Day, y=Connectedness, color=Year, group=ID)) +
    geom_line(alpha=0.25) +
    geom_point(alpha=0.35) +
    scale_y_continuous(breaks=1:10) +
    scale_x_continuous(breaks=unique(data.connectedness$Day)) +
    scale_color_colorblind() +
    theme_minimal()
) %>% ggplotly()

g_proximity <- (
    ggplot(c_jitter %>% filter(!is.na(Housing.F20)),
            aes(x=Day, y=Connectedness, color=Housing.F20, group=ID)) +
    geom_line(alpha=0.15) +
    geom_point(alpha=0.35) +
    scale_y_continuous(breaks=1:10) +
    scale_x_continuous(breaks=unique(data.connectedness$Day)) +
    scale_color_brewer(palette="Spectral") +
    theme_minimal()
) %>% ggplotly()

g_enrollment <- (
    ggplot(c_jitter %>% filter(!is.na(Enrolled.F20)),
           aes(x=Day, y=Connectedness, color=Enrolled.F20, shape=Enrolled.F20, group=ID)) +
    geom_line(alpha=0.25) +
    geom_point(alpha=0.35) +
    scale_y_continuous(breaks=1:10) +
    scale_x_continuous(breaks=unique(data.connectedness$Day)) +
    scale_color_colorblind() +
    theme_minimal()
) %>% ggplotly()

# https://stackoverflow.com/questions/36227482/r-ggplot-dodging-geom-lines
# https://ggplot2.tidyverse.org/reference/scale_brewer.html
```

```{r}
# subplot(g_year, g_proximity, g_enrollment, shareX=T, nrows=3)
g_year
g_proximity
g_enrollment
```