---
title: "Mapping Moose"
description: "Movements of the Herd"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

```{r echo=F}
library(tidyverse)
library(rnaturalearth)
library(tigris)  # gives state data
library(sf)
library(ggthemes)
library(plotly)
library(gganimate)  # also requires transformr
```

We want to animate people's locations over the past few months.

```{r}
t <- "20201216_104624"
data.locations.conn <- readRDS(paste("data/locations.conn_", t, ".rds", sep=""))
```

```{r results='hide'}
world <- ne_countries(scale = "medium", returnclass = "sf")
states <- states(class = "sf")
```

There are two similar approaches. The software library plotly offers a pretty nice way to turn ggplots into interactive visualizations which can can animate through time. Unfortunately, it currently appears to be a limitation that the animation slider's breakpoints must be evenly spaced.^[https://stackoverflow.com/questions/65334985/how-to-customize-plotly-r-ggplot2-animation-breakpoints-frame-spacing-duration] This is undesirable for us, since our waypoints and epochs are not evenly-spaced. Using day of the year seems misleading in this case, and a full English description of the epoch/waypoint is too long to fit, so we use an abbreviated Epoch/Waypoint label. Refer to the table below for the date/season itself.

```{r layout="l-body-outset"}
g <- ggplot(world) +
    geom_sf(fill="white") +
    geom_sf(data=data.locations.conn, aes(frame=TimeEW, ids=ID, color=Connectedness)) +
    scale_color_distiller(palette="PuBu", direction=1) +  # https://ggplot2.tidyverse.org/reference/scale_brewer.html
    theme_tufte()

p <- ggplotly(g) %>%
    animation_opts(redraw=F, easing="sin-out", frame=800) %>%
    animation_slider(
        currentvalue=list(font=list(color="black"), suffix=" (see below)")
    )

p
```

* t1 is before Spring Break (< 3/7)
* t2 is Spring Break week 1 (3/7 - 3/15)
* t2.5 is Spring Break week 2 (3/16 - 3/22)
* t3 is remainder of semester (3/23 - 5/6)
* t4 is Summer (5/7 - Aug)
* t5 is Fall semester, before break (8/31 - 11/21)
* t6 is remainder of Fall semester (11/22 - 2020)

If we sacrifice the interactiveness, a more true-to-time visualization is given by ``gganimate``.

```{r}
day2season <- function(d) {
    d <- as.numeric(d)
    if (d < epoch_days["2"])   return("Spring Semester, before Break")
    if (d < epoch_days["2.5"]) return("Spring Break Week 1")
    if (d < epoch_days["3"])   return("Spring Break Week 2")
    if (d < epoch_days["4"])   return("Rest of Spring")
    if (d < epoch_days["5"])   return("Summer")
    if (d < epoch_days["6"])   return("Fall Semester, before Break")
    if (d <= epoch_days["7"])  return("Rest of 2020")
    return(NA)
}
day2date <- function(d) {
    return(format(as.Date(toString(round(d)), format = "%j", origin="12-31-2019"), "%b %d"))
}

g2 <- ggplot(world) +
    geom_sf(fill="white") +
    geom_sf(data=data.locations.conn, aes(color=Connectedness)) +
    scale_color_distiller(palette="PuBu", direction=1) +
    theme_tufte() +
    labs(title = "{day2date(frame_time)} \t~ \t{day2season(frame_time)}") +
    transition_time(Day) +
    ease_aes("sine-out")

animate(g2, fps=10, nframes=260, end_pause=10)  # Requests an fps which divides 100
```


